---
title: "Using `pointblank` in ad hoc analysis"
author: "Katie Masiello"
date: "2/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Say you just received a new project and you've just gotten your hands on the first batch of data.  Let's go through some steps to explore this data with the help of `pointblank`

```{r packages, include = FALSE}
library(tidyverse)
library(pointblank)
```

### Bring in the data 
We're going to bring in data from the FAA Wildlife Strike Database with data from 1990 to 2018. 
```{r}
strikes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv")

```

There are many ways to get a feel for the data, including `summary(strikes)` or `skimr::skim(strikes)`
```{r}
summary(strikes)
skimr::skim(strikes)
```

Pointblank offers the `scan_data()` function to provide a comprehensive view of the data. (this can take some time to run... be patient, or consider running this as a background job) 
```{r}
scan_data(strikes)
```

